Environmental Market Design for Large-Scale Marine Conservation

### 2019-11-20 docs/Markets_Conservation_final.tex
Updated MS for resubmission

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these loses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we design and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that cross-country transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific Island Nations that collectively manages tuna purse seine fishing in its members' waters\cite{havice_2013,aqorau_2018} (Fig. \ref{fig:PNA_map}). These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP \cite{ffa_2017b}. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by January 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how a market can be designed to incentivize conservation.

@@

To examine how market design incentivizes or punishes large-scale conservation, we develop a 10-country spatial bioeconomic model that mirrors the strategies and spatial connections among the nine PNA nations and the high seas. Countries 1 to 9 represent the PNA countries, which operate under a vessel-day scheme where vessel-days are capped for each country and closely tracked. ``Country'' 10 represents the high seas, where fishing days are unregulated and determined by prevailing economic conditions. We examine the effects of large-scale conservation in a single country (Country 1) under markets with and without trading between countries. In all cases, we solve the bioeconomic model for the equilibrium vessel-day price, fishing effort redistribution across countries, and fish stock that would be expected to occur in the market. We quantify the change in revenue to Country 1 and compare each scenario to a benchmark scenario without any conservation action. We simulate these outcomes across a range of reserve sizes and assumptions about within-country stock movement (see the Methods section for details).

We first simulate a fishery where trading between countries is not allowed. This represents the status quo of any nation unilaterally engaging in large-scale conservation. Intuitively, we find that a spatial closure in Country 1 will always result in a loss in revenues to Country 1 (Fig. \ref{fig:PNA_model}a). Higher within-country stock movement ($\theta$ = 0 implies no within-country movement and $\theta = 1$ implies that fish are well-mixed within the fishing season) allows vessels to harvest the stock within the remaining open area, lowering the cost to Country 1. Even for a highly mobile stock where fish can move in and out of the reserve, a spatial closure reduces the amount of biomass that is available for harvest in the conserving country (\emph{i.e.} biomass outside the reserve), which reduces vessels' willingness to pay to fish in such waters (Fig. S1). When countries cannot trade, the costs of conservation are incurred by Country 1, but the benefits are received by the eight other countries (revenues increase between 0\% and 6\% each; Fig. S2) and the high seas. This benchmark simulation highlights the misalignment of incentives, where a conservation-minded country incurs large costs and provides public benefits to other nations with no mechanism for recouping these benefits.

How would trading between countries change these results? We simulate the same fishery, but now allow for vessel-days to be traded across countries (as is the case for the PNA). As before, a closure in Country 1 lowers the value of vessel-days in that country (because the fishable area shrinks). But increased biomass in other countries causes their vessel day prices to increase. As a result, vessel-days from Country 1 are traded to Countries 2 to 9, until prices equalize across countries (Fig. S3). Under this market design, revenue losses to Country 1 are less than 1\%, compared to the base case with no reserve (Fig. \ref{fig:PNA_model}b; note that the vertical axis now only ranges from 0 to 0.8\% instead of from 0 to 100\% as in Fig. \ref{fig:PNA_model}a). This shows that, with trading, the relative revenue drop will always be smaller than the relative effort drop, and the opposite is observed when there is no trading (Fig. S4). Overall, this shows that 88\% to 99\% of the costs of conservation can be avoided if markets are designed to allow trading (Fig. \ref{fig:PNA_model}c).

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ to implement the Phoenix Islands Protected Area (PIPA), effectively displacing all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat\cite{hernndez_2019}, PIPA may provide important recruitment and biomass benefits to adjacent waters. We combine vessel-level tracking data \cite{kroodsma_2018} and country-level license revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) \cite{ffa_2017} to quantify the displacement of vessel-days and the likely costs of conservation. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 254 ``non-displaced vessels'' never fished in PIPA waters but fished within PNA waters. We use the vessel-level tracking data to calculate vessel-days in the same way that the PNA does (see Methods for details). We present descriptive statistics on the redistribution of fishing activity before and after the implementation of PIPA.

Consistent with our model's prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig. \ref{fig:fishing_raster_diff}). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}a-b). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}a). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Figs. S6 and S7). The reduction in effort in Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicts.

The decrease of vessel-days in Kiribati could be alternatively (or jointly) explained by changes in oceanographic conditions that drive the distribution of target species and fishing vessels. For example, as El Ni\~no events develop, tuna species are known to shift longitudinally across PNA waters, causing vessels to redistribute \cite{aqorau_2018,hanich2018unraveling}. However, the aggregate decrease that we observe for Kiribati in Fig. \ref{fig:empirical}a is driven by the relocation of vessels that historically fished within PIPA, not by the entire fleet. Oceanographic conditions should be influencing the entire fleet whereas the closure of PIPA should have a stronger impact on vessels that used to fish there. Thus, the large drop in effort within Kiribati for the displaced vessels (relative to the non-displaced vessels) is consistent with the argument that PIPA displaced these vessels to waters outside of Kiribati. One could argue that differences in vessel characteristics between displaced and non-displaced vessels may influence their ability to redistribute or take advantage of different oceanographic conditions in Kiribati. It should be noted that, on average, displaced vessels have smaller crew sizes, more engine power, are larger than non-displaced vessels, fished more in the PNA in 2014, and are more likely to be registered to the Republic of Korea (Tables S2 and S3, Fig. S5). Thus, another alternative explanation for the observed changes in fishing patterns is a change in international relations between Kiribati (or other PNA countries) and the flag country of a particular vessel, although we are not aware of any occurring between 2012 and 2018. Ruling out alternative explanations for observed effects is an important component of careful empirical evaluation of MPAs \cite{ferraro2018causal}: as such, our results should not be interpreted as the direct causal impacts of PIPA but are better viewed as patterns that are consistent with the predictions of our theoretical model.

As predicted by our model, the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's water without large revenue losses. Kiribati's reported revenue increased from US\$127.3 million in 2014 to US\$148.8 million in 2015, before decreasing to US\$118.3 million in 2016 (Fig. \ref{fig:empirical}c). The increase and subsequent decrease in revenues matches the vessel-day patterns observed for Kiribati from 2014 to 2016 (Fig. \ref{fig:empirical}d). But, critically, the drop in revenue in 2016 (20\%) is smaller than the drop in vessel-days (35\%). This supports a key prediction of our model: with trading, the relative revenue drop will be smaller than the relative effort drop but, without trading, the opposite relationship would hold (Fig. S4). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Fig. S8), despite the PIPA closure.

@@

Our findings may help inform management and implementation of existing and upcoming MPAs in the PNA. In January 2020, Palau will close nearly 80\% of its EEZ to commercial fishing to create the Palau National Marine Sanctuary (PNMS): the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). Vessel-tracking data (2012 to 2018) shows that, on average, the proposed PNMS boundaries have historically contained 86 $\pm$ 5.30\% ($\pm$ 1SD) of longline vessel-days (non-tradable) and 91.3 $\pm$ 5.03\% ($\pm$ 1SD) of purse seine vessel-days (tradable) in Palauan waters (Fig. S10-S11).

While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from the current 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. A 100\% biomass-based allocation rule means that fishing rights will be distributed between nations based on the proportion of total biomass within their waters, regardless of historical fishing intensity. Furthermore, an effort-based allocation would reward undesirable behavior by granting more fishing rights to a country that fishes the most. This is a model prediction that cannot be empirically tested in the PNA context because the allocation rule has not been experimentally modified. Nevertheless, the PNA countries have already shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may also have become pioneers in effective large-scale marine conservation in a market-based setting.

@@

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. We use AIS data provided by Global Fishing Watch \cite{kroodsma_2018} to track 318 tuna purse seiners that fished within the PNA. For every georeferenced position, we observe the time spent (defined as the time since the last position), and whether the vessel was actively fishing versus only transiting. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation, the remaining 254 ``non-displaced vessels'' never fished in PIPA waters. Our dataset contains more than 37 million geo-referenced positions for these 318 tuna purse seiners. We use these data to calculate vessel-days (the metric used by the PNA), and to track the spatial redistribution of displaced vessels. A comparison of vessel characteristics between displaced and non-displaced vessels is presented in Tables S2-S3 and Figure S5.

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell. We use a 1-degree grid as a reasonable compromise between higher resolutions that would result in a more granular but noisy footprint, and the simple estimation of vessel-days at the EEZ-level (as in Fig S6, but represented spatially). This process results in 14 gridded datasets of fishing hours (7 years, for two groups). For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group, and normalize the value of each grid cell by diving it by the largest within-group absolute change:

@@

We obtained information on revenues from the Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report.  Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA \cite{ffa_2017}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:empirical}c; Fig. S8 - S9). For countries in the PNA, these revenues show a combination of vessel-day license fees as well as joint-venture operations.

@@

\subsection{Data availability}

@@

\subsection{Code availability}

@@

\bibliography{references}
\bibliographystyle{nature}

@@

@@

\section{Author information}

Reprints and permissions information is available at www.nature.com/reprints. The authors declare no competing interests. Correspondence and requests for materials should be addressed to juancarlos@ucsb.edu.

\subsection{Competing interests}

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these losses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we simulate and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a bioeconomic model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation under this market-based setting. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific Island Nations that collectively manages tuna purse seine fishing in its members' waters \cite{havice_2013,aqorau_2018} (Fig. \ref{fig:PNA_map}). These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP \cite{ffa_2017b}. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by January 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how an environmental market can be designed to incentivize conservation.

@@

To examine how market design incentivizes or hinders large-scale conservation, we develop a 10-country spatial bioeconomic model that mirrors the strategies and spatial connections among the nine PNA nations and the high seas. Countries 1 to 9 represent the PNA countries, which operate under a vessel-day scheme where vessel-days are capped for each country and closely tracked. ``Country'' 10 represents the high seas, where fishing days are unregulated and determined by prevailing economic conditions. We examine the effects of large-scale conservation in a single country (Country 1) under markets with and without trading between countries. In all cases, we solve the bioeconomic model for the equilibrium vessel-day price, fishing effort redistribution across countries, and fish stock that would be expected to occur in the market. We quantify the change in revenue to Country 1 and compare each scenario to a benchmark scenario without any conservation action. We simulate these outcomes across a range of reserve sizes and assumptions about within-country stock movement (see the Methods section for details).

We first simulate a fishery where trading between countries is not allowed. This represents the status quo of any nation unilaterally engaging in large-scale conservation. Intuitively, we find that a spatial closure in Country 1 will always result in a loss in revenues to Country 1 (Fig. \ref{fig:PNA_model}a). Higher within-country stock movement ($\theta$ = 0 implies no within-country movement and $\theta = 1$ implies that fish are well-mixed within the fishing season) allows vessels to harvest the stock within the remaining open area, lowering the cost to Country 1. Even for a highly mobile stock where fish can move in and out of the reserve, a spatial closure reduces the amount of biomass that is available for harvest in the conserving country (\emph{i.e.} biomass outside the reserve), which reduces vessels' willingness to pay to fish in such waters (Supplementary Figure 1). When countries cannot trade, the costs of conservation are incurred by Country 1, but the benefits are received by the eight other countries (revenues increase between 0\% and 6\% each; Supplementary Figure 2) and the high seas. This benchmark simulation highlights the misalignment of incentives, where a conservation-minded country incurs large costs and provides public benefits to other nations with no mechanism for recouping these benefits.

How would trading between countries change these results? We simulate the same fishery, but now allow for vessel-days to be traded across countries (as is the case for the PNA). As before, a closure in Country 1 lowers the value of vessel-days in that country (because the fishable area shrinks). But increased biomass in other countries causes their vessel day prices to increase. As a result, vessel-days from Country 1 are traded to Countries 2 to 9, until prices equalize across countries (Supplementary Figure 3). Under this market design, revenue losses to Country 1 are less than 1\%, compared to the base case with no reserve (Fig. \ref{fig:PNA_model}b; note that the vertical axis now only ranges from 0 to 0.8\% instead of from 0 to 100\% as in Fig. \ref{fig:PNA_model}a). This shows that, with trading, the relative revenue drop will always be smaller than the relative effort drop, and the opposite is observed when there is no trading (Supplementary Figure 4). Overall, this shows that 88\% to 99\% of the costs of conservation can be avoided if markets are designed to allow trading (Fig. \ref{fig:PNA_model}c).

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ to implement the Phoenix Islands Protected Area (PIPA), effectively displacing all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat \cite{hernndez_2019}, PIPA may provide important recruitment and biomass benefits to adjacent waters. We combine vessel-level tracking data \cite{kroodsma_2018} and country-level license revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) \cite{ffa_2017} to quantify the displacement of vessel-days and the likely costs of conservation. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 254 ``non-displaced vessels'' never fished in PIPA waters but fished within PNA waters. We use the vessel-level tracking data to calculate vessel-days in the same way that the PNA does (see Methods for details). We present descriptive statistics on the redistribution of fishing activity before and after the implementation of PIPA.

Consistent with our model's prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig. \ref{fig:fishing_raster_diff}). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}a-b). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}a). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Extended Data Figures 1 - 2). The reduction in effort in Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicts.

The decrease of vessel-days in Kiribati could be alternatively (or jointly) explained by changes in oceanographic conditions that drive the distribution of target species and fishing vessels. For example, as El Ni\~no events develop, tuna species are known to shift longitudinally across PNA waters, causing vessels to redistribute \cite{aqorau_2018,hanich2018unraveling}. However, the aggregate decrease that we observe for Kiribati in Fig. \ref{fig:empirical}a is driven by the relocation of vessels that historically fished within PIPA, not by the entire fleet. Oceanographic conditions should be influencing the entire fleet whereas the closure of PIPA should have a stronger impact on vessels that used to fish there. Thus, the large drop in effort within Kiribati for the displaced vessels (relative to the non-displaced vessels) is consistent with the argument that PIPA displaced these vessels to waters outside of Kiribati. One could argue that differences in vessel characteristics between displaced and non-displaced vessels may influence their ability to redistribute or take advantage of different oceanographic conditions in Kiribati. It should be noted that, on average, displaced vessels have smaller crew sizes, more engine power, are larger than non-displaced vessels, fished more in the PNA in 2014, and are more likely to be registered to the Republic of Korea (Supplementary Tables 2-3 and Supplementary Figure 5). Thus, another alternative explanation for the observed changes in fishing patterns is a change in international relations between Kiribati (or other PNA countries) and the flag country of a particular vessel, although we are not aware of any occurring between 2012 and 2018. Ruling out alternative explanations for observed effects is an important component of careful empirical evaluation of MPAs \cite{ferraro2018causal}: as such, our results should not be interpreted as the direct causal impacts of PIPA but are better viewed as patterns that are consistent with the predictions of our theoretical model.

As predicted by our model, the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's water without large revenue losses. Kiribati's reported revenue increased from US\$127.3 million in 2014 to US\$148.8 million in 2015, before decreasing to US\$118.3 million in 2016 (Fig. \ref{fig:empirical}c). The increase and subsequent decrease in revenues matches the vessel-day patterns observed for Kiribati from 2014 to 2016 (Fig. \ref{fig:empirical}d). But, critically, the drop in revenue in 2016 (20\%) is smaller than the drop in vessel-days (35\%). This supports a key prediction of our model: with trading, the relative revenue drop will be smaller than the relative effort drop but, without trading, the opposite relationship would hold (Supplementary Figure 4). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Supplementary Figure 6), despite the PIPA closure.

@@

Our findings may help inform management and implementation of existing and upcoming MPAs in the PNA. In January 2020, Palau will close nearly 80\% of its EEZ to commercial fishing to create the Palau National Marine Sanctuary (PNMS): the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). Vessel-tracking data (2012 to 2018) shows that, on average, the proposed PNMS boundaries have historically contained 86 $\pm$ 5.30\% ($\pm$ 1SD) of longline vessel-days (non-tradable) and 91.3 $\pm$ 5.03\% ($\pm$ 1SD) of purse seine vessel-days (tradable) in Palauan waters (Extended Data Figures 3 - 4).

While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from the current 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. A 100\% biomass-based allocation rule means that fishing rights will be distributed between nations based on the proportion of total biomass within their waters, regardless of historical fishing intensity. In contrast, an effort-based allocation would reward undesirable behavior by granting more fishing rights to a country that fishes the most. This is a model prediction that cannot be empirically tested in the PNA context because the allocation rule has not been experimentally modified. Nevertheless, the PNA countries have already shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may also have become pioneers in effective large-scale marine conservation in a market-based setting.

@@

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. We use AIS data provided by Global Fishing Watch \cite{kroodsma_2018} to track 318 tuna purse seiners that fished within the PNA. For every georeferenced position, we observe the time spent (defined as the time since the last position), and whether the vessel was actively fishing versus only transiting. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation, the remaining 254 ``non-displaced vessels'' never fished in PIPA waters. Our dataset contains more than 37 million geo-referenced positions for these 318 tuna purse seiners. We use these data to calculate vessel-days (the metric used by the PNA), and to track the spatial redistribution of displaced vessels. A comparison of vessel characteristics between displaced and non-displaced vessels is presented in Supplementary Tables 2-3 and Supplementary Figure 5.

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell. We use a 1-degree grid as a reasonable compromise between higher resolutions that would result in a more granular but noisy footprint, and the simple estimation of vessel-days at the EEZ-level (as in Extended Data Figure 1, but represented spatially). This process results in 14 gridded datasets of fishing hours (7 years, for two groups). For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group, and normalize the value of each grid cell by diving it by the largest within-group absolute change:

@@

We obtained information on revenues from the Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report.  Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA \cite{ffa_2017}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:empirical}c; Supplementary Figures 6 - 7). For countries in the PNA, these revenues show a combination of vessel-day license fees as well as joint-venture operations.

@@

\section{Data availability}

@@

\section{Code availability}

@@

%\bibliography{references}
%\bibliographystyle{nature}

\begin{thebibliography}{10}

\bibitem{oleary_2016}
O'Leary, B.~C., Winther-Janson, M., Bainbridge, J.~M., Aitken, J., Hawkins,
  J.~P., and Roberts, C.~M.
\newblock {\em Conservation letters}{ \bf 9}(6), 398--404 nov  (2016).

\bibitem{dinerstein_2019}
Dinerstein, E., Vynne, C., Sala, E., Joshi, A.~R., Fernando, S., Lovejoy,
  T.~E., Mayorga, J., Olson, D., Asner, G.~P., Baillie, J. E.~M., Burgess,
  N.~D., Burkart, K., Noss, R.~F., Zhang, Y.~P., Baccini, A., Birch, T., Hahn,
  N., Joppa, L.~N., and Wikramanayake, E.
\newblock {\em Sci. Adv.}{ \bf 5}(4), eaaw2869 apr  (2019).

\bibitem{sala_2017}
Sala, E. and Giakoumi, S.
\newblock {\em {ICES} Journal of Marine Science}{ \bf } (2017).

\bibitem{sala_2018}
Sala, E., Lubchenco, J., Grorud-Colvert, K., Novelli, C., Roberts, C., and
  Sumaila, U.~R.
\newblock {\em Marine Policy}{ \bf 91}(2), 11--13 may  (2018).

\bibitem{mcleod_2019}
Mcleod, E., Bruton-Adams, M., Förster, J., Franco, C., Gaines, G., Gorong, B.,
  James, R., Posing-Kulwaum, G., Tara, M., and Terk, E.
\newblock {\em Frontiers in Marine Science}{ \bf } (2019).

\bibitem{ramesh_2019}
Ramesh, N., Rising, J.~A., and Oremus, K.~L.
\newblock {\em Science}{ \bf 364}(6446), 1192--1196 jun  (2019).

\bibitem{hernndez_2019}
Hernández, C.~M., Witting, J., Willis, C., Thorrold, S.~R., Llopiz, J.~K., and
  Rotjan, R.~D.
\newblock {\em Scientific reports}{ \bf 9}(1), 10772 dec  (2019).

\bibitem{agardy_2018}
Agardy, T.
\newblock {\em {ICES} Journal of Marine Science}{ \bf 75}(3), 1183--1185 may
  (2018).

\bibitem{havice_2010}
Havice, E.
\newblock {\em Marine Policy}{ \bf 34}(5), 979--987 sep  (2010).

\bibitem{havice_2013}
Havice, E.
\newblock {\em Marine Policy}{ \bf 42}, 259--267 nov  (2013).

\bibitem{aqorau_2018}
Aqorau, T., Bell, J., and Kittinger, J.~N.
\newblock {\em Science}{ \bf 361}(6408), 1208.2--1209 sep  (2018).

\bibitem{ffa_2017b}
FFA.
\newblock Technical report, Pacific Islands Forum Fisheries Agency,  (2017).

\bibitem{libecap_1989}
Libecap, G.~D.
\newblock {\em Journal of Institutional and Theoretical Economics}{ \bf
  145}(1), 6--24 mar  (1989).

\bibitem{mccauley_2016}
{McCauley}, D.~J., Woods, P., Sullivan, B., Bergman, B., Jablonicky, C., Roan,
  A., Hirshfield, M., Boerder, K., and Worm, B.
\newblock {\em Science}{ \bf 351}(6278), 1148--1150 mar  (2016).

\bibitem{mcdermott_2018}
{McDermott}, G.~R., Meng, K.~C., {McDonald}, G.~G., and Costello, C.~J.
\newblock {\em Proc Natl Acad Sci {USA}}{ \bf } aug  (2018).

\bibitem{kroodsma_2018}
Kroodsma, D.~A., Mayorga, J., Hochberg, T., Miller, N.~A., Boerder, K.,
  Ferretti, F., Wilson, A., Bergman, B., White, T.~D., Block, B.~A., Woods, P.,
  Sullivan, B., Costello, C., and Worm, B.
\newblock {\em Science}{ \bf 359}(6378), 904--908 feb  (2018).

\bibitem{ffa_2017}
Terawasi, P. and Reid, C.
\newblock Technical report, Pacific Islands Forum Fisheries Agency, Honaira,
  Solomon Islands,  (2017).

\bibitem{hanich2018unraveling}
Hanich, Q., Rotjan, R., Aqorau, T., Bailey, M., Campbell, B., Gray, N., Gruby,
  R., Hampton, J., Ota, Y., Parris, H., et~al.
\newblock {\em Proceedings of the National Academy of Sciences}{ \bf 115}(52),
  E12122--E12123 (2018).

\bibitem{ferraro2018causal}
Ferraro, P.~J., Sanchirico, J.~N., and Smith, M.~D.
\newblock {\em Proceedings of the National Academy of Sciences}{ \bf },
  201805563 (2018).

\bibitem{crespo_2019}
Crespo, G.~O., Dunn, D.~C., Gianni, M., Gjerde, K., Wright, G., and Halpin,
  P.~N.
\newblock {\em Nature Ecology \& Evolution}{ \bf 3}(9), 1273--1276 sep  (2019).

\bibitem{rosegrant_1994}
Rosegrant, M.~W. and Binswanger, H.~P.
\newblock {\em World development}{ \bf 22}(11), 1613--1625 nov  (1994).

\bibitem{hutton_2003}
Hutton, J.~M. and Leader-Williams, N.
\newblock {\em Oryx}{ \bf 37}(02) apr  (2003).

\bibitem{sandbrook_2019}
Sandbrook, C., Fisher, J., Holmes, G., Luque-Lora, R., and Keane, A.
\newblock {\em Nat Sustain}{ \bf 2}(4), 316--323 apr  (2019).

\bibitem{roberts_2017}
Roberts, C.~M., O'Leary, B.~C., {McCauley}, D.~J., Cury, P.~M., Duarte, C.~M.,
  Lubchenco, J., Pauly, D., S√°enz-Arroyo, A., Sumaila, U.~R., Wilson,
  R.~W., Worm, B., and Castilla, J.~C.
\newblock {\em Proc Natl Acad Sci {USA}}{ \bf 114}(24), 6167--6175 jun  (2017).

\bibitem{ban_2019}
Ban, N.~C., Gurney, G.~G., Marshall, N.~A., Whitney, C.~K., Mills, M., Gelcich,
  S., Bennett, N.~J., Meehan, M.~C., Butler, C., Ban, S., Tran, T.~C., Cox,
  M.~E., and Breslow, S.~J.
\newblock {\em Nature Sustainability}{ \bf 2}(6), 524--532 jun  (2019).

\bibitem{costello_2016}
Costello, C., Ovando, D., Clavelle, T., Strauss, C.~K., Hilborn, R., Melnychuk,
  M.~C., Branch, T.~A., Gaines, S.~D., Szuwalski, C.~S., Cabral, R.~B., Rader,
  D.~N., and Leland, A.
\newblock {\em Proceedings of the National Academy of Sciences of the United
  States of America}{ \bf 113}(18), 5125--5129 may  (2016).

\bibitem{Hagrannsoknir2014}
Hagrannsoknir.
\newblock Technical report, PNA Office, available at: www.pnatuna.com/sites/
  default/files/Independent Review of the PNA Purse Seine VDS- Final Report_0.pdf,  (2014).

\bibitem{pna_2016}
PNA.
\newblock Technical report, Parties to the Nauru Agreement,  oct  (2016).

\bibitem{rcore_2019}
{R Core Team}.
\newblock {\em R: A Language and Environment for Statistical Computing}.
\newblock R Foundation for Statistical Computing, Vienna, Austria,  (2019).

\bibitem{yeeting2018stabilising}
Yeeting, A.~D., Weikard, H.-P., Bailey, M., Ram-Bidesi, V., and Bush, S.~R.
\newblock {\em Regional Environmental Change}{ \bf 18}(3), 885--897 (2018).

\end{thebibliography}

@@

Correspondence and requests for materials should be addressed to juancarlos@ucsb.edu.

@@

\section{Competing interests}
